### 2015-01-06 doc/thesis/evaluation.tex
Minor improvements to the text about typing LuaSocket and random case studies

We could type most of the members in the \texttt{socket} module,
which is the C core.
However, this module includes some functions such as
\texttt{socket.try}, \texttt{socket.protect}, and \texttt{socket.skip}
that have a dynamic behavior.
This means that the return type of these functions depend on the
type of the values that we use to call them.

@@

The function \texttt{ftp.get} downloads data from a given URL,
and it is difficult to type because its return type depend on
the type of its parameters.
Its type is either \texttt{(string) -> (string)?} or
\texttt{(url\string_argument|url\string_path) -> (number)?}.
The types \texttt{url\string_argument} and \texttt{url\string_path}
are table types that describe the URL that should be used to
download data.
Both table types include the same fields, except one:
while the first includes the mandatory field \texttt{argument},
of type \texttt{string}, the latter includes the mandatory field
\texttt{path}, also of type \texttt{string}.

The function \texttt{http.request} downloads data from a given URL,
and it is difficult to type because its return type depend on
the type of its parameters.
Its type is either
\texttt{(string, string?) -> (string, number, \{string : string\}, number)?} or
\texttt{(url\string_request) -> (number, number, \{string : string\}, number)?}.
The type \texttt{url\string_request} is a table type containing
details about the downloading URL.

@@

The modules \texttt{mime} and \texttt{ltn12} have a strong connection.
The \texttt{mime} module offers low-level and high-level filters
that apply and remove some text encodings.
The low-level filters are written in C, while the high-level filters
use the function \texttt{ltn12.filter.cycle} along with the low level
filters to create standard filters.
Even though we could type all the members of the \texttt{mime} module,
the function \texttt{ltn12.filter.cycle} is the only member of the
\texttt{ltn12} module that we could not give a precise type.
This function is difficult to type because the \texttt{mime} low-level
filters do not have a standard API.

@@

It is difficult to type the interface of the \texttt{http-digest} module
because it is an extension to the \texttt{http} module from LuaSocket.
The \texttt{http-digest} module only exports the function
\texttt{http-digest.request}, which extends the function
\texttt{http.request} with MD5 authentication.
Like \texttt{http.request}, the return type of \texttt{http-digest.request}
also depends on the type of its parameters.
Its type is either
\texttt{(string) -> (string, number, \{string : string\})?} or
\texttt{(url\string_request) -> (number, number, \{string : string\})?}.

@@

However, we hit some limitations of our type system to rewrite this module.
First, it uses the function \texttt{getmetatable} to get a table and
checks whether this table has the field \texttt{\string_\string_type} or not.
We could not give a precise type to \texttt{getmetatable}, so we used the dynamic
type \texttt{any} as its return type, and this generates a warning that we could not
remove.
Second, the module returns a metatable that extends \texttt{\string_\string_call}
with \texttt{typical.type}, that is, we can use the module itself as a function,
though it is a table.
Our type system still does not support metatables, so we did not extend our version
of the \texttt{typical} module to support \texttt{\string_\string_call}.
Third, the module uses \texttt{ipairs} to iterate over an array of functions,
but our type system also has limited support to \texttt{ipairs}, and generates
a warning when we try to use the indexed value inside the \texttt{for} body.
We removed this warning using the numeric \texttt{for} to perform the same loop.

>>>>

We could type most of the members in the \texttt{socket} module,
which is the C core.
However, this module includes some functions such as
\texttt{socket.try}, \texttt{socket.protect}, and \texttt{socket.skip}
that have a dynamic behavior.
This means that the return type of these functions depend on the
type of the values that we use to call them.
For instance, \texttt{socket.skip(0, nil, "hello")} returns the
tuple \texttt{(nil, "hello")}, \texttt{socket.skip(1, nil, "hello")}
returns only the string \texttt{"hello"}, and
\texttt{socket.skip(-1, nil, "hello")} returns the tuple
\texttt{(-1, nil, "hello")}.
Passing a negative number to \texttt{socket.skip} can be
dangerous, as it returns anything that might be in the stack.
For instance, \texttt{f = socket.skip(-2)} assigns \texttt{socket.skip}
to \texttt{f}.
Our type system cannot handle the type of negative numbers, as
this requires the refinement types from hybrid type checking \citep{flanagan2006htc}.

@@

The function \texttt{ftp.get} downloads data from a given URL,
and it is difficult to type because its return type depends on
the type of its input parameter.
More precisely, \texttt{ftp.get(url)} returns the tuple
\texttt{(string) | (nil, string)} if \texttt{url} is a string,
and it returns the tuple \texttt{(number) | (nil, string)} if
\texttt{url} is a table.

The function \texttt{http.request} downloads data from a given URL,
and it is difficult to type because its return type depends on
the types of its input parameters.
More precisely, \texttt{http.request(url, body)} returns the tuple
\texttt{(string, number, \{string:string\}, number) | (nil, string)}
if \texttt{url} is a string and \texttt{body} is another string or \texttt{nil},
but it returns the tuple
\texttt{(number, number, \{string:string\}, number) | (nil, string)}
if \texttt{url} is a table and \texttt{body} is \texttt{nil}.

@@

The modules \texttt{mime} and \texttt{ltn12} have a strong connection.
The \texttt{mime} module offers low-level and high-level filters
that apply and remove some text encodings.
The low-level filters are written in C, while the high-level filters
use the function \texttt{ltn12.filter.cycle} along with the low level
filters to create standard filters.

Even though we could type all the members of the \texttt{mime} module,
the function \texttt{ltn12.filter.cycle} is the only member of the
\texttt{ltn12} module that we could not give a precise type.
This function is difficult to type because the \texttt{mime} low-level
filters are polymorphic.

@@

It is difficult to type the interface of the \texttt{http-digest} module
because it is an extension to the \texttt{http} module from LuaSocket.
The \texttt{http-digest} module only exports the function
\texttt{http-digest.request}, which extends the function
\texttt{http.request} with MD5 authentication.
Like \texttt{http.request}, the return type of \texttt{http-digest.request}
also depends on the types of its input parameters.

@@

However, we hit some limitations of our type system to rewrite this module.
First, it uses the function \texttt{getmetatable} to get a table and
checks whether this table has the field \texttt{\string_\string_type} or not.
We could not give a precise type to \texttt{getmetatable}, so we used the dynamic
type \texttt{any} as its return type, and this generates a warning.
Second, the module returns a metatable that extends \texttt{\string_\string_call}
with \texttt{typical.type}, that is, we can use the module itself as a function,
though it is a table.
Our type system still does not support metatables, so we did not extend our version
of the \texttt{typical} module to support \texttt{\string_\string_call}.
Third, the module uses \texttt{ipairs} to iterate over an array of functions,
but our type system also has limited support to \texttt{ipairs}, and generates
a warning when we try to use the indexed value inside the \texttt{for} body.
We removed this warning using the numeric \texttt{for} to perform the same loop.
